A Decision Procedure for (Co)datatypes in SMT Solvers

### 2015-02-23 tex/body.tex
more

First, we define the set of interpretations of the theory $\thD$,
which requires additional terminology concerning $\mu$-terms.
Given a $\mu$-term $t$ with subterm $u$,
the \emph{expansion of $u$ with respect to $t$} is the $\mu$-term $\expand{t}{u}^\emptyset$,
abbreviated to
$\expand{t}{u}$, as returned by the following recursive function:
%where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
%\const{C}( \vec u )$ is the subterm of $t$ that binds this occurrence of
%variable $x$:
\[\begin{array}{r@{}c@{}l}
\expand{t}{x}^B & {} \,=\, {} &
\textstyle\relax{\begin{cases}
    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u}^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if this $x \notin B$ is
    bound by $\MU x.\; \const{C}( \vec u )$ in $t$}  \\[-\jot]
    x & \text{otherwise}
\end{cases}} ~ \\
\expand{t}{\MU x.\; \const{C}( \vec u )}^B & = &
\textstyle\begin{cases}
    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u }^{\smash{B \,\cup\, \{ x \}}}\bigr)} & \text{if }x \not\in B \\[-\jot]
    x & \text{otherwise}
\end{cases}
\end{array}\]
The recursion will eventually terminate because each recursion adds one bound
variable to $B$ and there are finitely many distinct bound variables in a $\mu$-term.
Intuitively, the expansion of a subterm is a stand-alone
$\mu$-term that denotes the same value as the original subterm---e.g.,
$\expand{\MU x.\: \const{C}(\MU y.\: \const{D}(x))}{\MU y.\: \const{D}(x)} =
 \MU y.\: \const{D}(\MU x.\: \const{C}(y))$.
%
%\rem{TODO: add small examples}
%\rem{TODO: explain why terminates}

>>>>

First, we define the set of interpretations of the theory $\thD$,
which requires additional terminology concerning $\mu$-terms.
Given a $\mu$-term $t$ with subterm $u$,
the \emph{expansion of $u$ with respect to $t$} is the $\mu$-term $\expand{t}{u}^\emptyset$,
abbreviated to
$\expand{t}{u}$, as returned by the following recursive function:
%where $x \leadsto \MU x.\; \const{C}( \vec u ) \in t$ indicates that $\MU x.\;
%\const{C}( \vec u )$ is the subterm of $t$ that binds this occurrence of
%variable $x$:
\[\begin{array}{@{}r@{}c@{}l@{}}
\expand{t}{x}^B & {} \,=\, {} &
\textstyle\relax{\begin{cases}
    x & \text{if $x \in B$}   \\[-\jot]
    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u}^{\smash{B \,\uplus\, \{ x \}}}\bigr)} & \text{if this occurrence of $x$ is
    bound by $\MU x.\; \const{C}( \vec u )$ in $t$}
\end{cases}} ~ \\
\expand{t}{\MU x.\; \const{C}( \vec u )}^B & = &
\textstyle\begin{cases}
    x & \text{if $x \in B$} \\[-\jot]
    \MU x.\; \const{C}\smash{\bigl( \expand{t}{ \vec u }^{\smash{B \,\uplus\, \{ x \}}}\bigr)} & \text{otherwise}
\end{cases}
\end{array}\]
The recursion will eventually terminate because each recursion adds one bound
variable to $B$ and there are finitely many distinct bound variables in a $\mu$-term.
Intuitively, the expansion of a subterm is a stand-alone
$\mu$-term that denotes the same value as the original subterm---e.g.,
$\expand{\MU x.\: \const{C}(\MU y.\: \const{D}(x))}{\MU y.\: \const{D}(x)} =
 \MU y.\: \const{D}(\MU x.\: \const{C}(y))$.
%
%\rem{TODO: add small examples}
%\rem{TODO: explain why terminates}
